---
title: "Demo - Detroit K-12 Education Network Map"
author: "Sankofa Consulting, LLC"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
library(igraph)
library(ggraph)
library(networkD3)
library(visNetwork)
library(htmlwidgets)
library(pander)
library(reactable)
library(scales)
```

## Purpose
This project maps the web of connections among organizations supporting K-12 education in Detroit to show how ideas, resources, and support flow across the community. By highlighting connections, this will help organizations see where they fit in, how they can contribute, and where bridging differences can open doors to stronger partnerships, greater visibility, and collective impact to improve education.

The data presented on this page were collected through an online survey shared with members of Detroit's K-12 education ecosystem. These data will be supplemented with qualitative insights generated through interviews and focus groups in an effort to understand the breadth and depth of connections across the network. This work was funded by the Skillman Foundation and conducted by Sankofa Consulting.


```{r Load & Prep Data, include=FALSE}

# Format system date
current_date <- Sys.Date()

formatted_date <- format(current_date, format = "%B %d, %Y")

# Load cleaned survey dataset
skillman_survey <- read_csv('/Users/Christian/Desktop/Sankofa/Skillman/Survey/Survey Results 2_19 Cleaned.csv')


# Renaming response organizations to align with survey naming conventions
skillman_survey$resp_org[skillman_survey$resp_org == "Michigan Association of Public School Academies"] <- "Michigan Association of Public School Academies (MAPSA)"
skillman_survey$resp_org[skillman_survey$resp_org == "Education Policy Innovation Collaborative - MSU"] <- "Education Policy Innovation Collaborative (EPIC) - Michigan State University"
skillman_survey$resp_org[skillman_survey$resp_org == "Detroit PAL"] <- "Detroit Police Athletic League"
skillman_survey$resp_org[skillman_survey$resp_org == "Detroit Employment Solutions Corporation"] <- "Detroit Employment Solutions Corporation (DESC) / Detroit at Work"
skillman_survey$resp_org[skillman_survey$resp_org == "CENTER FOR BLACK EDUCATOR DEVELOPMENT"] <- "Center for Black Educator Development"
skillman_survey$resp_org[skillman_survey$resp_org == "citizens research council of MI"] <- "Citizens Research Council of Michigan"
skillman_survey$resp_org[skillman_survey$resp_org == "EdTrust Midwest"] <- "EdTrust-Midwest"
skillman_survey$resp_org[skillman_survey$resp_org == "Skillman Foundation"] <- "The Skillman Foundation"
skillman_survey$resp_org[skillman_survey$resp_org == "DPSCD-Athletics"] <- "Detroit Public Schools Community District (DPSCD)"
skillman_survey$resp_org[skillman_survey$resp_org == "East side community network"] <- "Eastside Community Network"
skillman_survey$resp_org[skillman_survey$resp_org == "Michigan Education Justice Coalition"] <- "Michigan Education Justice Coalition (MEJC)"
skillman_survey$resp_org[skillman_survey$resp_org == "Data Driven Detroit"] <- "Data Driven Detroit (D3)"
skillman_survey$resp_org[skillman_survey$resp_org == "482forward"] <- "482Forward"
skillman_survey$resp_org[skillman_survey$resp_org == "Cody Rouge Community Action Alliance"] <- "Cody Rouge Neighborhood Alliance"
skillman_survey$resp_org[skillman_survey$resp_org == "Detroit Regional Chamber / Detroit Promise"] <- "Detroit Regional Chamber"
skillman_survey$resp_org[skillman_survey$resp_org == "Communities in Schools MI"] <- "Communities in Schools Michigan"
skillman_survey$resp_org[skillman_survey$resp_org == "The Michigan Association of Secondary School Principals"] <- "Michigan Association of Secondary School Principals (MASSP)"
skillman_survey$resp_org[skillman_survey$resp_org == "Bridge Detroit"] <- "Bridge Detroit/ Bridge Michigan"
skillman_survey$resp_org[skillman_survey$resp_org == "Detroit Public Media/Detroit PBS"] <- "Detroit Public Media / Detroit PBS"
skillman_survey$resp_org[skillman_survey$resp_org == "College Career and Beyond / Midnight Golf Program"] <- "College Career and Beyond"
skillman_survey$resp_org[skillman_survey$resp_org == "Detroit Partnership for Education Equity & Research (Detroit PEER)"] <- "Detroit Partnership for Education Equity & Research (PEER) - Wayne State University"
skillman_survey$resp_org[skillman_survey$resp_org == "EPIC - MSU"] <- "Education Policy Innovation Collaborative (EPIC) - Michigan State University"
skillman_survey$resp_org[skillman_survey$resp_org == "Crain's Detroit Business"] <- "Crain’s Business Detroit"
skillman_survey$resp_org[skillman_survey$resp_org == "Institute for Law and Organizing"] <- "Detroit Institute for Law and Organizing"


### Creating dataframe with level 1 interaction

# Subsetting to level 1 interaction questions only
skillman_l1 <- skillman_survey %>% 
  select(resp_no, resp_org, q4a:q6f, q10a:q12h, q16a:q18m, q22a:q24j, q28a:q30l) %>% 
  select(-contains("_none"),-contains("_other"))


## Creating a loop to append all columns

# Starting off with the first pair of columns
level_1_edgelist <- skillman_l1 %>% select(resp_org, q4a) %>% 
  mutate() %>% rename(from = resp_org,
                      to = q4a)


# Now looping with remainder
for (col in 4:ncol(skillman_l1)){
  new_level_1_edgelist <- cbind(skillman_l1[,2],skillman_l1[,col])
  colnames(new_level_1_edgelist)[c(1,2)] <- c("from", "to")
  level_1_edgelist <- rbind(level_1_edgelist, new_level_1_edgelist)
}


# Dropping non-existent connections
level_1_edgelist <- na.omit(level_1_edgelist)


# Removing self-edges (orgs who selected themselves)
level_1_edgelist <- level_1_edgelist %>% filter(from != to)


# Deduplicating from and to (non-directional, so only unique pairs matter)
level_1_edgelist <- level_1_edgelist %>% arrange(to) %>%
  mutate(from_id = as.numeric(as.factor(level_1_edgelist$from)),
         to_id = as.numeric(as.factor(level_1_edgelist$to)))
  
level_1_edgelist <- level_1_edgelist %>% rowwise() %>% 
  mutate(id_1 = min(from_id, to_id),
         id_2 = max(from_id, to_id))

level_1_edgelist <- level_1_edgelist %>%
  distinct(id_1, id_2, .keep_all = TRUE) %>% 
  select(from, to)

level_1_edgelist <- level_1_edgelist %>%
  distinct(from, to)


### Repeating to get a list of level 2 connections

# Subsetting to level 2 connection questions
skillman_l2 <- skillman_survey %>% 
  select(resp_no, resp_org, q7a:q9f, q13a:q15h, q19a:q21m, q25a:q27j, q31a:q33l) %>% 
  select(-contains("_none"),-contains("_other"))


## Creating a loop to append all columns

# Starting off with the first pair of columns
level_2_edgelist <- skillman_l2 %>% select(resp_org, q7a) %>% 
  mutate() %>% rename(from = resp_org,
                      to = q7a)


# Now looping with remainder
for (col in 4:ncol(skillman_l2)){
  new_level_2_edgelist <- cbind(skillman_l2[,2],skillman_l2[,col])
  colnames(new_level_2_edgelist)[c(1,2)] <- c("from", "to")
  level_2_edgelist <- rbind(level_2_edgelist, new_level_2_edgelist)
}


# Dropping non-existent connections
level_2_edgelist <- na.omit(level_2_edgelist)


# Removing self-edges (orgs who selected themselves)
level_2_edgelist <- level_2_edgelist %>% filter(from != to)


# Deduplicating from and to (non-directional, so only unique pairs matter)
level_2_edgelist <- level_2_edgelist %>% arrange(to) %>%
  mutate(from_id = as.numeric(as.factor(level_2_edgelist$from)),
         to_id = as.numeric(as.factor(level_2_edgelist$to)))

level_2_edgelist <- level_2_edgelist %>% rowwise() %>% 
  mutate(id_1 = min(from_id, to_id),
         id_2 = max(from_id, to_id))

level_2_edgelist <- level_2_edgelist %>%
  distinct(id_1, id_2, .keep_all = TRUE) %>% 
  select(from, to) %>% mutate(level = 2)


level_2_edgelist <- level_2_edgelist %>%
  distinct(from, to, level)


## Combining into one edgelist with a column for level of interaction
edgelist <- level_1_edgelist %>%  left_join(level_2_edgelist, by = c("from", "to"))

edgelist$level[is.na(edgelist$level)] <- 1


## Creating a comprehensive node list

# Starting from list of orgs in survey
survey_orgs <- read_csv('/Users/Christian/Desktop/Sankofa/Skillman/Survey/Survey Org List.csv')


# Now pulling list from to
response_orgs <- edgelist %>%
  select(from) %>%
  distinct() %>%
  mutate() %>% 
  rename(Organization = from)


# Combining and creating an ID
org_list <- survey_orgs %>% 
  select(Organization) %>% 
  rbind(response_orgs) %>% 
  distinct() %>% 
  arrange(Organization)
  
org_list$org_id <- 1:nrow(org_list)

# Zero-indexing id column
org_list <- org_list %>% mutate(across(c(org_id), ~  . - 1))



# Pulling in sector
org_list <- survey_orgs %>% select(Organization, Sector) %>% 
  right_join(org_list, by = c("Organization"))


# Labeling sector for response orgs not on survey - start with survey question
response_org_sector <- skillman_survey %>% 
  select(resp_org, resp_org_cat) %>% 
  distinct() %>% 
  arrange(resp_org)


# Adding to org list
org_list <- org_list %>% full_join(response_org_sector, by = c("Organization" = "resp_org"))


# Replacing missing values with sector question from survey
org_list$Sector <- ifelse(is.na(org_list$Sector), org_list$resp_org_cat, org_list$Sector)


# Manually labeling missing values
org_list$Sector[org_list$Organization == "826michigan"] <- "Youth Serving Organizations"
org_list$Sector[org_list$Organization == "College for Creative Studies, Precollege and Continuing Studies Dept."] <- "Institutes of Higher Education"
org_list$Sector[org_list$Organization == "North End Woodward Community Coalition (NEWCC)"] <- "Family and Community Organizations"
org_list$Sector[org_list$Organization == "Teach Plus Michigan"] <- "Policy and Advocacy Organizations"
org_list$Sector[org_list$Organization == "Detroit Regional Chamber"] <- "Business and Industry Organizations"
org_list$Sector[org_list$Organization == "Detroit Partnership for Education Equity & Research (PEER) - Wayne State University"] <- "Research and Data Organizations"
org_list$Sector[org_list$Organization == "Urban Neighborhood Initiatives"] <- "Family and Community Organizations"
org_list$Sector[org_list$Organization == "Detroit Institute for Law and Organizing"] <- "Policy and Advocacy Organizations"


# Dropping survey question column
org_list <- org_list %>% select(-resp_org_cat) %>% 
  arrange(Organization)


# Dropping extra row - one org shows up in two sectors
org_list <- org_list %>% filter(!(Organization == "Detroit Regional Workforce Partnership" & Sector == "Workforce Pathways and Training"))


## Pulling org ID into edgelist as from and to ID
edges <- edgelist %>% 
  left_join(org_list, by = c("from" = "Organization")) %>% 
  select(from, org_id, to, level) %>% 
  mutate() %>% rename(from_id = org_id)

edges <- edges %>% 
  left_join(org_list, by = c("to" = "Organization")) %>% 
  select(from, from_id, to, org_id, level) %>% 
  mutate() %>% rename(to_id = org_id)


# Creating a zero-indexed row ID for the edgelist
edges$row_id <- 1:nrow(edges)
edges <- edges %>% mutate(across(c(row_id), ~  . - 1)) %>% 
  select(row_id, from, from_id, to, to_id, level)



## Generating a few summary network statistics

# Responses - 65
response_orgs %>% n_distinct()

# Network size - 174 organizations
nrow(org_list)

# Density (connections / possible connections) - 6.33%
nrow(edges)/(nrow(org_list)*(nrow(org_list)-1))

# Degree centrality
from <- edges %>% select(from) %>% mutate() %>% rename(Organization = from)

to <- edges %>% select(to) %>% mutate() %>% rename(Organization = to)

all_edges <- rbind(from, to)

edge_count <- all_edges %>% 
  group_by(Organization) %>%
  summarise(edges = n())


# Betweenness centrality
edges_for_graph <- edges %>% select(from, to)

edges_graph <- graph_from_edgelist(as.matrix(edges_for_graph), directed = F)

V(edges_graph)$betweenness <- betweenness(edges_graph, directed = F)

edge_count$betweenness <- betweenness(edges_graph, directed = F)


# Closeness
closeness(edges_graph)

edge_count$closeness <- closeness(edges_graph)


# Adding the stats to the full org list

org_list <- org_list %>% left_join(edge_count)

org_list$edges <- ifelse(is.na(org_list$edges), 0, org_list$edges)

org_list$betweenness <- ifelse(is.na(org_list$betweenness), 0, org_list$betweenness)

org_list$closeness <- ifelse(is.na(org_list$closeness), 0, org_list$closeness)



## Prepping for visNetwork package

skillman_edges <- edges %>%
  select(from_id, to_id, level) %>%
  mutate() %>% rename(from = from_id,
                      to = to_id,
                      width = level)

skillman_edges$dashes <- ifelse(skillman_edges$width == 1, TRUE, FALSE)

skillman_vertices <- org_list %>% 
  mutate() %>% rename(id = org_id,
                      label = Organization,
                      group = Sector)

policy_priorities <- skillman_survey %>% 
  select(resp_org, q3a:q3h) %>% 
  group_by(resp_org) %>%
  filter(row_number() == 1)

policy_priorities$q3b[policy_priorities$q3b == "Educator Recruitment, Retention, and Support"] <- "Educator Recruitment / Retention / Support"

policy_priorities <- policy_priorities %>%
  unite(prior_pol, q3a, q3b, q3c, q3d, q3e, q3f, q3g, q3h, sep = ",", na.rm = TRUE)

skillman_vertices <- skillman_vertices %>% left_join(policy_priorities, by = c("label" = "resp_org"))


## Sector-Level Analysis

sector_from <- edges %>%
  left_join(skillman_vertices, by = c("from_id" = "id")) %>% 
  select(from, from_id, group, to, to_id, level) %>% 
  mutate() %>% rename(from_sector = group)

sector_to <- sector_from %>% 
  left_join(skillman_vertices, by = c("to_id" = "id")) %>% 
  select(from, from_id, from_sector, to, to_id, group, level) %>% 
  mutate() %>% rename(to_sector = group)

sector_edgelist <- sector_to %>% select(from_sector, to_sector, level) %>% 
  group_by(from_sector, to_sector, level) %>% 
  summarise(edges = n())

# Removing self-edges (within-sector connections)
sector_edgelist <- sector_edgelist %>% filter(from_sector != to_sector)

sector_vertices <- skillman_vertices %>% group_by(group) %>% 
  summarise(orgs = n())

sector_vertices$id <- 1:nrow(sector_vertices)

# Zero-indexing id column
sector_vertices <- sector_vertices %>% mutate(across(c(id), ~  . - 1))

sector_vertices <- sector_vertices %>% 
  mutate() %>% rename(sector = group,
                      value = orgs) %>% 
  select(sector, id, value)

sector_edges_from_id <- sector_edgelist %>%
  left_join(sector_vertices, by = c("from_sector" = "sector")) %>% 
  select(from_sector, id, to_sector, level, edges) %>% 
  mutate() %>% rename(from_id = id)

sector_edges_to_id <- sector_edges_from_id %>% 
  left_join(sector_vertices, by = c("to_sector" = "sector")) %>% 
  select(from_sector, from_id, to_sector, id, level, edges) %>% 
  mutate() %>% rename(to_id = id)

sector_edges <- sector_edges_to_id %>% ungroup() %>%
  select(from_id, to_id, level, edges) %>% 
  mutate() %>% rename(width = edges)

sector_edges$dashes <- ifelse(sector_edges$level == 1, TRUE, FALSE)

sector_edges <- sector_edges %>%
  select(from_id, to_id, width, dashes)


# Deduplicating from and to (non-directional, so only unique pairs matter)
sector_edges <- sector_edges %>% rowwise() %>% 
  mutate(id_1 = min(from_id, to_id),
         id_2 = max(from_id, to_id))

sector_edges$joint_id <- paste(sector_edges$id_1, sector_edges$id_2)

sector_edges <- sector_edges %>% group_by(joint_id, dashes) %>% 
  summarise(width = sum(width))

sector_edges <- sector_edges %>%
  separate(joint_id, into = c("from_id", "to_id")) %>% 
  select(from_id, to_id, width, dashes)


## Flagging Across-/Within-Sector Interaction
edges_by_sector <- edges %>%
  left_join(skillman_vertices, by = c("from_id" = "id")) %>% 
  select(row_id, from, from_id, group, to, to_id, level) %>% 
  mutate() %>% rename(from_sector = group)

edges_by_sector <- edges_by_sector %>% 
  left_join(skillman_vertices, by = c("to_id" = "id")) %>% 
  select(row_id, from, from_id, from_sector, to, to_id, group, level) %>% 
  mutate() %>% rename(to_sector = group)

edges_by_sector$within_sector <- ifelse(edges_by_sector$from_sector == edges_by_sector$to_sector, 1, 0)

from_sector <- edges_by_sector %>% select(from, within_sector) %>% mutate() %>% rename(Organization = from)

to_sector <- edges_by_sector %>% select(to, within_sector) %>% mutate() %>% rename(Organization = to)

all_edges_sector <- rbind(from_sector, to_sector)

edge_count_sector <- all_edges_sector %>% 
  group_by(Organization) %>%
  summarise(within_sector = sum(within_sector))


skillman_vertices <- skillman_vertices %>% left_join(edge_count_sector, by = c("label" = "Organization"))

skillman_vertices$within_sector[is.na(skillman_vertices$within_sector)] <- 0


legend_nodes <- data.frame(
  label = c("Sector",
            "Youth Serving Organizations",
            "Family and Community Organizations",
            "K-12 Education Systems Member Organizations and Unions",
            "K-12 Education Systems",
            "Charter Authorizers",
            "Institutes of Higher Education",
            "City and County Government",
            "State Policymakers and Agencies",
            "Policy and Advocacy Organizations",
            "Investors and Foundations",
            "Business and Industry Organizations",
            "Workforce Pathways and Training",
            "Research and Data Organizations",
            "Technical Assistance Providers",
            "Media"), 
  color = c("white",
            "darkblue",
            "blue",
            "lightblue",
            "purple",
            "darkorchid",
            "orchid",
            "darkgreen",
            "green",
            "coral",
            "darkred",
            "red",
            "brown",
            "gold",
            "yellow",
            "deeppink"),
  shape = c("dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot"),
  size = c(30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30),
  font.size = c(30, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20)
)
```

## Survey Summary
The Detroit K–12 Education Ecosystem Survey was first distributed on October 21, 2025. As of `r formatted_date`, we have received responses from `r response_orgs %>% n_distinct()` unique organizations. Based on desk review and contributions from survey respondents, the known network includes `r nrow(org_list)` organizations. Across this network, there were a total of `r nrow(edges)` unique connections between any two organizations, yielding a network density of `r round(nrow(edges)/(nrow(org_list)*(nrow(org_list)-1))*100, digits = 2)`%. This means that there is a `r round(nrow(edges)/(nrow(org_list)*(nrow(org_list)-1))*100, digits = 2)`% chance that any two organizations interacted with one another.

## Sector Statistics
The organizations in the network were grouped by their roles into one of 15 categories, what we refer to as "sectors" on this page. These sectors include everything from public and charter K-12 education systems, youth and community organizations, to government entities, investors, and foundations - and many more. The tables below present summary statistics by sector.

### Connections
In the survey, we asked respondents to indicate whether they had interacted with each organization we identified as being a part of the network within the last year. These interactions were separated into two levels:

- **Level 1:** Any interaction, formal or informal.
- **Level 2:** Close interaction, through shared projects, coordination, or active collaboration.

Respondents were only asked to evaluate close interaction (**Level 2**) for organizations with whom they had indicated any interaction (**Level 1**). Simply put, having a **Level 2** interaction presupposes a **Level 1** interaction.

The table below presents counts of organizations by sector, their total number of connections (**Level 1** or **Level 2**), and the average number of connections per organization.

```{r Table of Connections by Sector, echo=FALSE}
# Let's collapse these variables in a pander table:
skill_vert_label <- skillman_vertices %>% mutate(Sector = group)

panderOptions('round',2)
skill_vert_label %>% select(Sector, id, edges, within_sector) %>%
  group_by(Sector) %>% 
  summarize(`Organizations` = n_distinct(id),
            `Total Connections` = sum(edges),
            `Connections per Organization` = round((sum(edges)/n_distinct(id)), digits = 1),
            `Percent within Sector` = round((sum(within_sector)/sum(edges))*100, digits = 1),
            .groups = "rowwise") %>% 
  arrange(-`Connections per Organization`) %>% 
  pander(style = "rmarkdown", caption = "Connections by Sector", split.tables = 150)
```


### Measures of Centrality
There are several metrics for centrality in network analysis. The most common is **degree centrality**, which is simply a count of all of the connections for a given node. This metric is presented in the table above as an average in the *Connections per Organization* column. This network is *non-directed*, meaning it does not matter whether a connection is coming from or going to a given organization - all that matters is that they are connected. So in this analysis, **degree centrality** is the sum of all unique *to* and *from* connections.

Another measure, **closeness**, calculates the relative centrality of a given node by considering its direct and indirect connections - this factors in degrees of separation from other nodes. An organization's **closeness** score depends both on how many organizations with which it is directly connected *and* how many organizations those connected organizations have

A third, common metric is **betweenness**, which measures the degree to which a given node connects other nodes in the network. An organization may not have many direct connections, but if it serves as an important bridge within the network, it will have a high **betweenness** score.

The table below presents the average values of these additional measures of centrality by sector.

```{r Table of Centrality by Sector, echo=FALSE}
# Let's collapse these variables in a pander table:
skill_vert_label <- skillman_vertices %>% mutate(Sector = group)

panderOptions('round',2)
skill_vert_label %>% select(Sector, id, edges, betweenness, closeness) %>%
  group_by(Sector) %>% 
  summarize(`Organizations` = n_distinct(id),
            `Avg. Degree Centrality` = round((sum(edges)/n_distinct(id)), digits = 1),
            `Avg. Closeness` = round((sum(closeness)/n_distinct(id))*1000, digits = 1),
            `Avg. Betweenness` = round((sum(betweenness)/n_distinct(id)), digits = 1),
            .groups = "rowwise") %>% 
  arrange(-`Avg. Degree Centrality`) %>% 
  pander(style = "rmarkdown", caption = "Centrality by Sector", split.tables = 150)
```

<br>
<br>

#### Top 5 Organizations by Degree Centrality
The table below presents the five organizations with the highest **degree centrality** score - *total number of connections* - within the Detroit K-12 education ecosystem.

```{r Top 5 Orgs by Degree Centrality, echo=FALSE}
# Let's collapse these variables in a pander table:
skill_vert_label <- skillman_vertices %>% mutate(Sector = group)

panderOptions('round',2)
skill_vert_label %>% select(label, Sector, edges, betweenness, closeness) %>%
  arrange(-edges) %>% 
  head(5) %>% 
  mutate(`Organization` = label,
            `Sector` = Sector,
            `Degree Centrality` = edges,
            `Closeness` = round((closeness)*1000, digits = 1),
            `Betweenness` = betweenness) %>% 
  arrange(-`Degree Centrality`) %>% 
  select(Organization, Sector, `Degree Centrality`, Closeness, Betweenness) %>% 
  pander(style = "rmarkdown", caption = "Top 5 Organizations by Degree Centrality", split.tables = 150)
```

<br>
<br>

#### Top 5 Organizations by Closeness
The table below presents the five organizations with the highest **closeness** score - *direct and indirect connections* - within the Detroit K-12 education ecosystem.

```{r Top 5 Orgs by Closeness, echo=FALSE}
# Let's collapse these variables in a pander table:
skill_vert_label <- skillman_vertices %>% mutate(Sector = group)

panderOptions('round',2)
skill_vert_label %>% select(label, Sector, edges, betweenness, closeness) %>%
  arrange(-closeness) %>% 
  head(5) %>% 
  mutate(`Organization` = label,
            `Sector` = Sector,
            `Degree Centrality` = edges,
            `Closeness` = round((closeness)*1000, digits = 1),
            `Betweenness` = betweenness) %>% 
  arrange(-`Closeness`) %>% 
  select(Organization, Sector, `Degree Centrality`, Closeness, Betweenness) %>% 
  pander(style = "rmarkdown", caption = "Top 5 Organizations by Closeness", split.tables = 150)
```

<br>
<br>

#### Top 5 Organizations by Betweenness
The table below presents the five organizations with the highest **betweenness** score - *how well these organizations serve as a bridge* - within the Detroit K-12 education ecosystem.

```{r Top 5 Orgs by Betweenness, echo=FALSE}
# Let's collapse these variables in a pander table:
skill_vert_label <- skillman_vertices %>% mutate(Sector = group)

panderOptions('round',2)
skill_vert_label %>% select(label, Sector, edges, betweenness, closeness) %>%
  arrange(-betweenness) %>% 
  head(5) %>% 
  mutate(`Organization` = label,
            `Sector` = Sector,
            `Degree Centrality` = edges,
            `Closeness` = round((closeness)*1000, digits = 1),
            `Betweenness` = betweenness) %>% 
  arrange(-`Betweenness`) %>% 
  select(Organization, Sector, `Degree Centrality`, Closeness, Betweenness) %>% 
  pander(style = "rmarkdown", caption = "Top 5 Organizations by Betweenness", split.tables = 150)
```



## Network Maps
The following maps provide a visual depiction of the organizations within and connections across Detroit's K-12 education ecosystem. There are a few things worth noting before engaging with the maps themselves.

First, they are created using a physics-based algorithm. Think of each node (organization) as a metal ball with a magnetic charge, repelling the other balls around it. Each edge (connection) serves as a spring, helping to overcome this magnetic charge and create a structure for the network.

Because of this physics-based approach, the orientation of the organizations within the network is somewhat random. Physical proximity on the map does not correlate with closeness of interaction, nor does distance indicate the lack of interaction.

Lastly, **Level 1** connections are depicted using *dashed* lines, while **Level 2** connections appear as *solid* lines.

### How to Use the Maps
These maps are interactive - they include the following functionality:

- Use a mouse or trackpad to pan across the map or zoom in and out.
- Click or hover over an organization to see its immediate connections.
- Use the dropdown list to select a specific organization and see its immediate connections.
- Use the secondary dropdown to select several nodes based on sector or policy priority.
- Use a mouse or trackpad to zoom in and out on the legend on the left-hand side.


### Sector-Level Network Map
The network map below displays connections at the sector level. The size of the nodes corresponds to the number of organizations in that sector. The width of the edges corresponds to the total number of connections between the two sectors. Click a node or use the dropdown menu to select a sector and see its immediate connections.

```{r Prep Tables for Network Map, include=FALSE}
skillman_vert_scale <- skillman_vertices %>% mutate() %>% 
  rename(value = edges) %>% 
  mutate(title = paste0("<p><b>", label, "</b><br>",value," Connections"))

sector_vertices <- sector_vertices %>%
  mutate() %>% rename(label = sector) %>% 
  mutate(group = label,
         title = paste0("<p><b>", label, "</b><br>",value," Connections")) %>% 
  select(label, group, id, value, title)

sector_edges <- sector_edges %>% 
  mutate(from = as.numeric(from_id),
         to = as.numeric(to_id)) %>% 
  select(from, to, width, dashes)

skillman_edges_2 <- skillman_edges %>% filter(dashes == FALSE)
```

```{r Network Map at Sector-Level, echo=FALSE}
visNetwork(sector_vertices, sector_edges) %>% 
  visOptions(width = 625,
             height = 600,
             highlightNearest = list(enabled = TRUE, hover = TRUE,
                                     degree = 1, labelOnly = FALSE),
             nodesIdSelection = list(enabled = TRUE, useLabels = TRUE,
                                     style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                                     main = "Select by Sector")) %>%
  visGroups(groupname = "Youth Serving Organizations", color = "darkblue") %>% 
  visGroups(groupname = "Family and Community Organizations", color = "blue") %>% 
  visGroups(groupname = "K-12 Education Systems Member Organizations and Unions", color = "lightblue") %>% 
  visGroups(groupname = "K-12 Education Systems", color = "purple") %>% 
  visGroups(groupname = "Charter Authorizers", color = "darkorchid") %>% 
  visGroups(groupname = "Institutes of Higher Education", color = "orchid") %>% 
  visGroups(groupname = "City and County Government", color = "darkgreen") %>% 
  visGroups(groupname = "State Policymakers and Agencies", color = "green") %>% 
  visGroups(groupname = "Policy and Advocacy Organizations", color = "coral") %>% 
  visGroups(groupname = "Investors and Foundations", color = "darkred") %>% 
  visGroups(groupname = "Business and Industry Organizations", color = "red") %>% 
  visGroups(groupname = "Workforce Pathways and Training", color = "brown") %>% 
  visGroups(groupname = "Research and Data Organizations", color = "gold") %>% 
  visGroups(groupname = "Technical Assistance Providers", color = "yellow") %>% 
  visGroups(groupname = "Media", color = "deeppink") %>%
  visPhysics(solver = "forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant = -40)) %>% 
  visNodes(shape = "dot",
           scaling = list(min = 20, max = 50),
           labelHighlightBold = TRUE)
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

### Network Map by Policy Priority
The network map below allows the user to select multiple organizations by policy priority (as stated on the survey). Using the policy priority dropdown list will enable users to identify the organizations aligned around a given policy. This can be used to take stock of existing coalitions and identify opportunities to partner with other organizations with similar goals.

```{r Network Map with Selection by Policy Priority, echo=FALSE}
visNetwork(skillman_vert_scale, skillman_edges) %>% 
  visOptions(width = 625,
             height = 600,
             highlightNearest = list(enabled = TRUE,
                                     degree = 1, labelOnly = FALSE),
             nodesIdSelection = list(enabled = TRUE, useLabels = TRUE,
                                     style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                                     main = "Select by Organization"),
             selectedBy = list(variable = "prior_pol",
                               highlight = FALSE,
                               multiple = TRUE,
                               style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                               main = "Select by Policy Priority")) %>%
  visGroups(groupname = "Youth Serving Organizations", color = "darkblue") %>% 
  visGroups(groupname = "Family and Community Organizations", color = "blue") %>% 
  visGroups(groupname = "K-12 Education Systems Member Organizations and Unions", color = "lightblue") %>% 
  visGroups(groupname = "K-12 Education Systems", color = "purple") %>% 
  visGroups(groupname = "Charter Authorizers", color = "darkorchid") %>% 
  visGroups(groupname = "Institutes of Higher Education", color = "orchid") %>% 
  visGroups(groupname = "City and County Government", color = "darkgreen") %>% 
  visGroups(groupname = "State Policymakers and Agencies", color = "green") %>% 
  visGroups(groupname = "Policy and Advocacy Organizations", color = "coral") %>% 
  visGroups(groupname = "Investors and Foundations", color = "darkred") %>% 
  visGroups(groupname = "Business and Industry Organizations", color = "red") %>% 
  visGroups(groupname = "Workforce Pathways and Training", color = "brown") %>% 
  visGroups(groupname = "Research and Data Organizations", color = "gold") %>% 
  visGroups(groupname = "Technical Assistance Providers", color = "yellow") %>% 
  visGroups(groupname = "Media", color = "deeppink") %>% 
  visLegend(addNodes = legend_nodes,
            useGroups = FALSE,
            width = 0.2,
            stepY = 125) %>%
  visPhysics(solver = "forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant = -20)) %>% 
    visNodes(shape = "dot",
           scaling = list(min = 20, max = 50),
           labelHighlightBold = TRUE)
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

### Network Map by Policy Priority - Level 2 Connections Only
The network map below functions similarly to the map above, but has been filtered to include only **Level 2** connections. The map allows the user to select multiple organizations by policy priority (as stated on the survey). Using the policy priority dropdown list will enable users to identify the organizations aligned around a given policy. This can be used to take stock of existing coalitions and identify opportunities to partner with other organizations with similar goals.

```{r Level 2 Network Map with Selection by Policy Priority, echo=FALSE}
visNetwork(skillman_vert_scale, skillman_edges_2) %>% 
  visOptions(width = 625,
             height = 600,
             highlightNearest = list(enabled = TRUE,
                                     degree = 1, labelOnly = FALSE),
             nodesIdSelection = list(enabled = TRUE, useLabels = TRUE,
                                     style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                                     main = "Select by Organization"),
             selectedBy = list(variable = "prior_pol",
                               highlight = FALSE,
                               multiple = TRUE,
                               style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                               main = "Select by Policy Priority")) %>%
  visGroups(groupname = "Youth Serving Organizations", color = "darkblue") %>% 
  visGroups(groupname = "Family and Community Organizations", color = "blue") %>% 
  visGroups(groupname = "K-12 Education Systems Member Organizations and Unions", color = "lightblue") %>% 
  visGroups(groupname = "K-12 Education Systems", color = "purple") %>% 
  visGroups(groupname = "Charter Authorizers", color = "darkorchid") %>% 
  visGroups(groupname = "Institutes of Higher Education", color = "orchid") %>% 
  visGroups(groupname = "City and County Government", color = "darkgreen") %>% 
  visGroups(groupname = "State Policymakers and Agencies", color = "green") %>% 
  visGroups(groupname = "Policy and Advocacy Organizations", color = "coral") %>% 
  visGroups(groupname = "Investors and Foundations", color = "darkred") %>% 
  visGroups(groupname = "Business and Industry Organizations", color = "red") %>% 
  visGroups(groupname = "Workforce Pathways and Training", color = "brown") %>% 
  visGroups(groupname = "Research and Data Organizations", color = "gold") %>% 
  visGroups(groupname = "Technical Assistance Providers", color = "yellow") %>% 
  visGroups(groupname = "Media", color = "deeppink") %>% 
  visLegend(addNodes = legend_nodes,
            useGroups = FALSE,
            width = 0.2,
            stepY = 125) %>%
  visPhysics(solver = "forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant = -20)) %>% 
  visNodes(shape = "dot",
           scaling = list(min = 20, max = 50))
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

## Interactive Tables
The following table can be sorted, searched, and reformatted...

### By Organization
Each row in this table represents an organization depicted in the network maps above. Click on the column headers to sort the rows (double click to sort in *descending* order). Use the search bars below the column headers to identify specific organizations, sectors, or policy priorities.

```{r Interactive Table Org, echo=FALSE}
skillman_vert_tab <- skillman_vertices[, c(1,2,7,4)]

skillman_vert_tab$prior_pol <-gsub(",", ", ", skillman_vert_tab$prior_pol)

reactable(
  skillman_vert_tab,
  striped = TRUE,
  columns = list(
    group = colDef(name = "Sector", defaultSortOrder = "asc"),
    label = colDef(name = "Organization"),
    edges = colDef(name = "Connections", filterable = FALSE),
    prior_pol = colDef(name = "Policy Priorities")
  ),
  defaultSorted = c("group"),
  filterable = TRUE,
  defaultPageSize = 10
)


```


### By Sector
Each row in this table represents an organization depicted in the network maps above, grouped and aggregated by their sector. Click on the column headers to sort the rows (double click to sort in *descending* order). Click on a row to expand it and show its member organizations, double-click to collapse it again.

```{r Interactive Table Sector, echo=FALSE}
skillman_vert_tab <- skillman_vertices[, c(2,1,7,4)]

skillman_vert_tab$prior_pol <-gsub(",", ", ", skillman_vert_tab$prior_pol)

reactable(
  skillman_vert_tab,
  groupBy = "group",
  striped = TRUE,
  columns = list(
    group = colDef(name = "Sector", defaultSortOrder = "asc", filterable = FALSE),
    label = colDef(name = "Organization"),
    edges = colDef(name = "Connections", aggregate = "mean", format = colFormat(digits = 1), filterable = FALSE),
    prior_pol = colDef(name = "Policy Priorities")
  ),
  defaultSorted = c("group"),
  defaultPageSize = 15
)


```

### By Policy Priority
Each row in this table represents an organization depicted in the network maps above, grouped and aggregated by their stated policy priorities. Click on the column headers to sort the rows (double click to sort in *descending* order). Click on a row to expand it and show its member organizations, double-click to collapse it again.

```{r Interactive Table Pol, echo=FALSE}
skillman_vert_pol <- skillman_vertices[, c(7,1,2,4)] %>% 
  separate_longer_delim(cols = prior_pol, delim = ",") %>% 
  mutate(prior_pol = replace_na(prior_pol, "None Listed"),
         prior_pol = replace(prior_pol, prior_pol == "", "None Listed"))

reactable(
  skillman_vert_pol,
  groupBy = "prior_pol",
  striped = TRUE,
  columns = list(
    group = colDef(name = "Sector"),
    label = colDef(name = "Organization"),
    edges = colDef(name = "Connections", aggregate = "mean", format = colFormat(digits = 1), filterable = FALSE),
    prior_pol = colDef(name = "Policy Priorities", defaultSortOrder = "asc")
  ),
  defaultSorted = c("prior_pol"),
  defaultPageSize = 10
)


```


## Skillman Grantees

```{r Load & Prep Grantee Data, include=FALSE}
# Load list of orgs by grantee status
orgs_by_grant_status <- read_csv('/Users/Christian/Desktop/Sankofa/Skillman/Survey/Grantee Org List.csv')


## Flagging Grantee-/Non-Grantee Interaction
edges_by_grantee <- edges %>%
  left_join(skillman_vertices, by = c("from_id" = "id")) %>% 
  select(row_id, from, from_id, group, to, to_id, level) %>% 
  mutate() %>% rename(from_sector = group) %>% 
  left_join(orgs_by_grant_status, by = c("from_id" = "id")) %>% 
  mutate() %>% rename(from_grantee = grantee)

edges_by_grantee <- edges_by_grantee %>% 
  left_join(skillman_vertices, by = c("to_id" = "id")) %>% 
  select(row_id, from, from_id, from_sector, from_grantee, to, to_id, group, level) %>% 
  mutate() %>% rename(to_sector = group) %>% 
  left_join(orgs_by_grant_status, by = c("to_id" = "id")) %>% 
  mutate() %>% rename(to_grantee = grantee) %>% 
  select(row_id, from, from_id, from_sector, from_grantee, to, to_id, to_sector, to_grantee, level)

edges_by_grantee <- edges_by_grantee %>% 
  mutate(grant_to_grant = case_when(from_grantee == "Yes" & to_grantee == "Yes" ~ 1,
                                    TRUE ~ 0))

from_grantee <- edges_by_grantee %>% select(from, grant_to_grant) %>% mutate() %>% rename(Organization = from)

to_grantee <- edges_by_grantee %>% select(to, grant_to_grant) %>% mutate() %>% rename(Organization = to)

all_edges_grantee <- rbind(from_grantee, to_grantee)

edge_count_grantee <- all_edges_grantee %>% 
  group_by(Organization) %>%
  summarise(grant_to_grant = sum(grant_to_grant))

skillman_vertices <- skillman_vertices %>% left_join(edge_count_grantee, by = c("label" = "Organization"))

skillman_vertices$grant_to_grant[is.na(skillman_vertices$grant_to_grant)] <- 0

skill_vert_label_grant <- skill_vert_label %>% left_join(orgs_by_grant_status, by = "id") %>% 
  select(-label.y) %>% 
  mutate() %>% 
  rename(label = label.x) %>% 
  left_join(edge_count_grantee, by = c("label" = "Organization"))

skill_vert_label_grant$grant_to_grant[is.na(skill_vert_label_grant$grant_to_grant)] <- 0

skill_vert_label_grant_only <- skill_vert_label_grant %>% filter(grantee == "Yes")

all_edges_g2g_only <- all_edges_grantee %>% filter(grant_to_grant == 1)

skillman_vertices_grant <- orgs_by_grant_status %>% select(id,grantee) %>% right_join(skillman_vertices) %>% select(label, group, id, edges, betweenness, closeness, prior_pol, within_sector, grant_to_grant, grantee)

skillman_vertices_grant$group[skillman_vertices_grant$grantee == "No"] <- "Non-Grantee"

skillman_vert_scale_grant <- skillman_vertices_grant %>% mutate() %>% 
  rename(value = edges) %>% 
  mutate(title = paste0("<p><b>", label, "</b><br>",value," Connections"))

skillman_vert_scale_grant_only <- skillman_vert_scale_grant %>% filter(grantee == "Yes")

skillman_edges_from_grant <- skillman_edges %>% 
  left_join(orgs_by_grant_status, by = c("from" = "id")) %>% 
  mutate() %>% 
  rename(from_grantee = grantee) %>% 
  select(-label)

skillman_edges_to_grant <- skillman_edges_from_grant %>% 
  left_join(orgs_by_grant_status, by = c("to" = "id")) %>% 
  mutate() %>% 
  rename(to_grantee = grantee) %>% 
  select(-label)

skillman_edges_grant_only <- skillman_edges_to_grant %>% 
  filter(from_grantee == "Yes" & to_grantee == "Yes")

legend_nodes_grant <- data.frame(
  label = c("Sector",
            "Youth Serving Organizations",
            "Family and Community Organizations",
            "K-12 Education Systems Member Organizations and Unions",
            "K-12 Education Systems",
            "Charter Authorizers",
            "Institutes of Higher Education",
            "City and County Government",
            "State Policymakers and Agencies",
            "Policy and Advocacy Organizations",
            "Investors and Foundations",
            "Business and Industry Organizations",
            "Workforce Pathways and Training",
            "Research and Data Organizations",
            "Technical Assistance Providers",
            "Media",
            "Non-Grantee"), 
  color = c("white",
            "darkblue",
            "blue",
            "lightblue",
            "purple",
            "darkorchid",
            "orchid",
            "darkgreen",
            "green",
            "coral",
            "darkred",
            "red",
            "brown",
            "gold",
            "yellow",
            "deeppink",
            "grey"),
  shape = c("dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot",
            "dot"),
  size = c(30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 30, 15),
  font.size = c(30, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 20, 10)
)

```

This section provides detailed network information for current Skillman Foundation grant partners. Within the network are `r skill_vert_label_grant_only %>% n_distinct()` unique grantee organizations. There were a total of `r nrow(skillman_edges_grant_only)` connections between grantees, representing `r round(nrow(skillman_edges_grant_only)/nrow(edges)*100, digits = 2)`% of all connections within the network. The network density for grantees was `r round(nrow(skillman_edges_grant_only)/(nrow(skill_vert_label_grant_only)*(nrow(skill_vert_label_grant_only)-1))*100, digits = 2)`%, compared to an overall network density of `r round(nrow(edges)/(nrow(org_list)*(nrow(org_list)-1))*100, digits = 2)`%. This means that the odds that any two grantees are connected are `r round((nrow(skillman_edges_grant_only)/(nrow(skill_vert_label_grant_only)*(nrow(skill_vert_label_grant_only)-1)))/(nrow(edges)/(nrow(org_list)*(nrow(org_list)-1))), digits = 1)`X higher than the odds that any two organizations within the network are connected.

The table below presents summary statistics for Skillman grantees, grouped by sector. These statistics include all connections for these organizations, including those with non-grantees. The "% w/ Other Grantees" column presents the percentage of all connections for grantees in that sector that were made with other grant partners.

```{r Table of Connections by Sector - Grantees, echo=FALSE}
# Let's collapse these variables in a pander table:
panderOptions('round',2)
skill_vert_label_grant_only %>% select(Sector, id, edges, within_sector, grant_to_grant) %>%
  group_by(Sector) %>% 
  summarize(`Organizations` = n_distinct(id),
            `Total Connections` = sum(edges),
            `Avg. Connections` = round((sum(edges)/n_distinct(id)), digits = 1),
            `% w/in Sector` = round((sum(within_sector)/sum(edges))*100, digits = 1),
            `% w/ Other Grantees` = round((sum(grant_to_grant)/sum(edges))*100, digits = 1),
            .groups = "rowwise") %>% 
  arrange(-`Organizations`) %>% 
  pander(style = "rmarkdown", caption = "Connections by Sector, Skillman Grant Partners", split.tables = 150)
```

<br>
<br>

### Grantee Centrality
The statistics below present the same measures of centrality discussed above, subset to Skillman grant partners only.

```{r Table of Centrality by Sector - Grantees, echo=FALSE}
panderOptions('round',2)
skill_vert_label_grant_only %>% select(Sector, id, edges, betweenness, closeness) %>%
  group_by(Sector) %>% 
  summarize(`Organizations` = n_distinct(id),
            `Avg. Degree Centrality` = round((sum(edges)/n_distinct(id)), digits = 1),
            `Avg. Closeness` = round((sum(closeness)/n_distinct(id))*1000, digits = 1),
            `Avg. Betweenness` = round((sum(betweenness)/n_distinct(id)), digits = 1),
            .groups = "rowwise") %>% 
  arrange(-`Avg. Degree Centrality`) %>% 
  pander(style = "rmarkdown", caption = "Centrality by Sector, Skillman Grant Partners", split.tables = 150)
```

<br>
<br>

#### Top 5 Grantees by Degree Centrality
The table below presents the five grantees with the highest **degree centrality** score - *total number of connections* - within the Detroit K-12 education ecosystem.

```{r Top 5 Orgs by Degree Centrality - Grantees, echo=FALSE}
panderOptions('round',2)
skill_vert_label_grant_only %>% select(label, Sector, edges, betweenness, closeness) %>%
  arrange(-edges) %>% 
  head(5) %>% 
  mutate(`Organization` = label,
            `Sector` = Sector,
            `Degree Centrality` = edges,
            `Closeness` = round((closeness)*1000, digits = 1),
            `Betweenness` = betweenness) %>% 
  arrange(-`Degree Centrality`) %>% 
  select(Organization, Sector, `Degree Centrality`, Closeness, Betweenness) %>% 
  pander(style = "rmarkdown", caption = "Top 5 Organizations by Degree Centrality, Skillman Grant Partners", split.tables = 150)
```

<br>
<br>

#### Top 5 Grantees by Closeness
The table below presents the five grantees with the highest **closeness** score - *direct and indirect connections* - within the Detroit K-12 education ecosystem.

```{r Top 5 Orgs by Closeness - Grantees, echo=FALSE}
panderOptions('round',2)
skill_vert_label_grant_only %>% select(label, Sector, edges, betweenness, closeness) %>%
  arrange(-closeness) %>% 
  head(5) %>% 
  mutate(`Organization` = label,
            `Sector` = Sector,
            `Degree Centrality` = edges,
            `Closeness` = round((closeness)*1000, digits = 1),
            `Betweenness` = betweenness) %>% 
  arrange(-`Closeness`) %>% 
  select(Organization, Sector, `Degree Centrality`, Closeness, Betweenness) %>% 
  pander(style = "rmarkdown", caption = "Top 5 Organizations by Closeness, Skillman Grant Partners", split.tables = 150)
```

<br>
<br>

#### Top 5 Grantees by Betweenness
The table below presents the five grantees with the highest **betweenness** score - *how well these organizations serve as a bridge* - within the Detroit K-12 education ecosystem.

```{r Top 5 Orgs by Betweenness - Grantees, echo=FALSE}
# Let's collapse these variables in a pander table:
skill_vert_label <- skillman_vertices %>% mutate(Sector = group)

panderOptions('round',2)
skill_vert_label %>% select(label, Sector, edges, betweenness, closeness) %>%
  arrange(-betweenness) %>% 
  head(5) %>% 
  mutate(`Organization` = label,
            `Sector` = Sector,
            `Degree Centrality` = edges,
            `Closeness` = round((closeness)*1000, digits = 1),
            `Betweenness` = betweenness) %>% 
  arrange(-`Betweenness`) %>% 
  select(Organization, Sector, `Degree Centrality`, Closeness, Betweenness) %>% 
  pander(style = "rmarkdown", caption = "Top 5 Organizations by Betweenness, Skillman Grant Partners", split.tables = 150)
```
<br>
<br>

### Grantee Network Maps
The map below presents all actors and connections (both *Level 1* and *Level 2*) within the network. Non-grant partners are presented in grey and scaled smaller, while grant partners are presented in colors corresponding to their sector, and scaled based on their total number of connections.

#### All Network Members
```{r Grantee Network Map with Selection by Policy Priority, echo=FALSE}
visNetwork(skillman_vert_scale_grant, skillman_edges) %>% 
  visOptions(width = 625,
             height = 600,
             highlightNearest = list(enabled = TRUE,
                                     degree = 1, labelOnly = FALSE),
             nodesIdSelection = list(enabled = TRUE, useLabels = TRUE,
                                     style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                                     main = "Select by Organization"),
             selectedBy = list(variable = "prior_pol",
                               highlight = FALSE,
                               multiple = TRUE,
                               style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                               main = "Select by Policy Priority")) %>%
  visGroups(groupname = "Youth Serving Organizations", color = "darkblue") %>% 
  visGroups(groupname = "Family and Community Organizations", color = "blue") %>% 
  visGroups(groupname = "K-12 Education Systems Member Organizations and Unions", color = "lightblue") %>% 
  visGroups(groupname = "K-12 Education Systems", color = "purple") %>% 
  visGroups(groupname = "Charter Authorizers", color = "darkorchid") %>% 
  visGroups(groupname = "Institutes of Higher Education", color = "orchid") %>% 
  visGroups(groupname = "City and County Government", color = "darkgreen") %>% 
  visGroups(groupname = "State Policymakers and Agencies", color = "green") %>% 
  visGroups(groupname = "Policy and Advocacy Organizations", color = "coral") %>% 
  visGroups(groupname = "Investors and Foundations", color = "darkred") %>% 
  visGroups(groupname = "Business and Industry Organizations", color = "red") %>% 
  visGroups(groupname = "Workforce Pathways and Training", color = "brown") %>% 
  visGroups(groupname = "Research and Data Organizations", color = "gold") %>% 
  visGroups(groupname = "Technical Assistance Providers", color = "yellow") %>% 
  visGroups(groupname = "Media", color = "deeppink") %>% 
  visGroups(groupname = "Non-Grantee", color = "grey") %>% 
  visLegend(addNodes = legend_nodes_grant,
            useGroups = FALSE,
            width = 0.2,
            stepY = 125) %>%
  visPhysics(solver = "forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant = -20)) %>% 
    visNodes(shape = "dot",
           scaling = list(min = 20, max = 50),
           labelHighlightBold = TRUE)
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

#### Grantees Only
The map below presents Skillman grant partners alone, and only the connections between Skillman grant partners.
```{r Grantee Only Network Map with Selection by Policy Priority, echo=FALSE}
visNetwork(skillman_vert_scale_grant_only, skillman_edges_grant_only) %>% 
  visOptions(width = 625,
             height = 600,
             highlightNearest = list(enabled = TRUE,
                                     degree = 1, labelOnly = FALSE),
             nodesIdSelection = list(enabled = TRUE, useLabels = TRUE,
                                     style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                                     main = "Select by Organization"),
             selectedBy = list(variable = "prior_pol",
                               highlight = FALSE,
                               multiple = TRUE,
                               style = 'width: 300px; height: 26px;
                                     background: #f8f8f8;
                                     color: darkblue;
                                     border:none;
                                     outline:none;',
                               main = "Select by Policy Priority")) %>%
  visGroups(groupname = "Youth Serving Organizations", color = "darkblue") %>% 
  visGroups(groupname = "Family and Community Organizations", color = "blue") %>% 
  visGroups(groupname = "K-12 Education Systems Member Organizations and Unions", color = "lightblue") %>% 
  visGroups(groupname = "K-12 Education Systems", color = "purple") %>% 
  visGroups(groupname = "Charter Authorizers", color = "darkorchid") %>% 
  visGroups(groupname = "Institutes of Higher Education", color = "orchid") %>% 
  visGroups(groupname = "City and County Government", color = "darkgreen") %>% 
  visGroups(groupname = "State Policymakers and Agencies", color = "green") %>% 
  visGroups(groupname = "Policy and Advocacy Organizations", color = "coral") %>% 
  visGroups(groupname = "Investors and Foundations", color = "darkred") %>% 
  visGroups(groupname = "Business and Industry Organizations", color = "red") %>% 
  visGroups(groupname = "Workforce Pathways and Training", color = "brown") %>% 
  visGroups(groupname = "Research and Data Organizations", color = "gold") %>% 
  visGroups(groupname = "Technical Assistance Providers", color = "yellow") %>% 
  visGroups(groupname = "Media", color = "deeppink") %>% 
  visLegend(addNodes = legend_nodes,
            useGroups = FALSE,
            width = 0.2,
            stepY = 125) %>%
  visPhysics(solver = "forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant = -20)) %>% 
    visNodes(shape = "dot",
           scaling = list(min = 20, max = 50),
           labelHighlightBold = TRUE)
```


